Simulation of Nano Particles in a Laser Trap

### 2017-02-22 doc/draft4.tex
updates

\lhead[\ ]{\ }
\chead[\ ]{\ }
\rhead[\ ]{\ }

\lfoot[\ ]{\ }
\cfoot[\ ]{\ }
\rfoot[\ ]{\ }

@@

\begin{document}
\title{Simulation of Nano Particles in a Laser Trap}
\author{Mathias H\"old, BSc.}
\date{2016}
\maketitle
\thispagestyle{empty}
\newpage
\tableofcontents
\newpage
\section{Introduction}
%something about computer science in general, optical tweezers and the need to study the
%systems on computers yada yada
In the year 1953, Metropolis et al.~\cite{Metropolis1953} used the MANIAC (Mathematical Analyzer Numerical Integrator And Computer) 
at the Los Alamos National Laboratories in the United States to
simulate particles in a liquid state to calculate the equation of state. This was the first publication in which scientists used a computer to
calculate numerical results based on a physical theory in a process that is to this day called \textit{Metropolis Monte Carlo}. To say that a lot has
changed in the field of computer simulation (or computers in general) would be a vast understatement.\\
Not only has the processing power of computers increased exponentially (Moore's law), which allows the simulation of a greater number of particles for
a longer period of time, but also new techniques have been developed, such as molecular dynamics. These techniques are not only used by physicists,
but also find applications in the fields of chemistry, biology, etc., making it possible to apply them to virtually anything. This thesis is an
attempt to apply methods of simulation to a field of active research: optical trapping.
Optical trapping is a technique that has been used and developed for over 40 years~\cite{Marago2013}. The idea for this goes back to the beginning of
the 20th century, where Lebedev~\cite{Lebedew1901}, Nichols and Hull~\cite{Nichols1901} demonstrated the existence of a force that light exerts on 
matter. Due to the technical limitations at that time, the idea was discarded until 1970, where Ashkin~\cite{Ashkin2000} used lasers to manipulate the
motion of neutral atoms and micrometer sized objects and thus brought the idea of optical trapping back to life.\\
With this advancement it was possible to develop methods for trapping and manipulating objects of different sizes ranging from the micrometer scale
down to the subnano scale. At first, this technique was used to either trap objects in the micrometer range (like cells) or in the subnano scale,
where lasers were used for the cooling of atoms. The range between those two, with objects at the scale of one to a hundred nanometer, was not
explored as much because neither downscaling the methods for micrometer sized objects nor upscaling the methods for atoms seemed feasible. In recent
years however, new developments in the field opened the possibilities to work with a variety of objects in the desired nanometer scale, such as metal
nanoparticles, nano tubes, quantum dots and much more. This non-invasive method of trapping objects allows very precise measurements of various
properties and also allows the cooling of nanoparticles towards their quantum mechanical ground state.\\
The small size of the objects used in the optical tweezer experiments make them ideal candidates for computer simulations. In this thesis, I will
model an optical tweezer setup, based on a real experiment, to investigate different temperatures in a non-equilibrium state.\\
The first chapter will be the introduction of the experiment and the introduction of the problem statement. The second chapter contains basic
information on the techniques used to model the experiment and simulate the dynamic properties. In the third chapter, I will present my findings which
are followed by the conclusions and further remarks.

@@

The particle fluctuates within the trap in all three spatial directions. These fluctuations can be approximated
such that they are decoupled, which means that they can be described by a one-dimensional Langevin equation:

@@

The one dimensional Langevin equation \eqref{eq:langevin} has to be modeled to include the different friction coefficients of the two heat baths. The

@@

\newpage
\section{Simulation}
%simulation techniques and general concept of the simulation itself, used methods and so on 
The problem at hand can be studied on an atomic level with the use of computer simulation. There is a variety of methods for computer simulations
that are widely used, one of which being molecular dynamics (MD) simulations.\\ 
The goal is to simulate the experiment depicted in Fig.~\ref{fig:setup} as accurately as possible. To achieve this, the setup has to be broken down
into individual pieces that can be modeled by using existing methods. The nano particle will be approximated by a cube of particles, sitting on a FCC
(face-centered cubic) lattice that interact via a Lennard-Jones potential. The laser will be broken down into its two main purposes, 
as it acts as a trapping device as well as a heat source for the nano particle. Surrounding the nano particle will be a gas chamber, 
that acts as a thermostat in the simulation, which will be modeled as a cubic box surrounding the nano particle. 
This box is filled with gas particles that interact with the nano particle via a soft-sphere potential and do not interact with one another.\\
The following section will give a brief overview of the concepts used to simulate every part of the experiment.

@@

Argon has a mass of m = $6.69 \times 10^{-26}$ kg, $\sigma = 3.4 \times 10^{-10}$ m and $\varepsilon = 1.65 \times 10^{-21}$ J.\\

@@

The force in the y and z direction can be calculated analogously.\\

@@

energy after this point is vanishingly small. To save computing time it is common practice to cut off the potential at a certain point and don't

@@

    p_2 &=& \{0.5*a,0.5*a,0\}\\
    p_3 &=& \{0.5*a,0,0.5*a\}\\
    p_4 &=& \{0,0.5*a,0.5*a\}

@@

The sum of those two equations yields

@@

The problem with these kind of algorithms is, that a target temperature has to be fixed which will be reached at some point. In order to simulate the

@@

Without any equilibrating mechanism the setup described by now would lead to the system heating up indefinetly, which is not desirable. Furthermore,

@@

    \item Randomly draw the number of particles that are created on a single side of the minimal volume of cells, $N_\text{fac}$, from the
        distribution 

@@

        desired temperature. This chosen number of particles is then equally distributed over the face of the cell on which they are created. The
        velocities of the created particles are drawn from two different random number distributions. The component of the particle perpendicular to

@@

        (where U is the interaction energy between the gas particle and the crystal) because it is possible that the inserted particle is 

@@

    \mathbf{v}_\text{tot} &=& \sum_{i=1}^N \mathbf{v}_i\\
    \mathbf{v}_\text{cut} &=& \frac{\mathbf{v}_\text{tot}}N\\
    \mathbf{v}_i^\text{new} &=& \mathbf{v}_i - \mathbf{v}_\text{cut}

@@

The barostat, as described previously, keeps the system from overheating. The barostat algorithm is applied at the same time as the eHEX algorithm,

@@

While the eHEX algorithm is pumping energy into the nano particle while the surrounding gas particles absorb excessive energy. This leads to the

@@

To check the integrity of the algortihm, a simple case needs to be tested. Such a simple case is $\Delta Q = 0$, i.e. no energy begin pumped into the

@@

    T_\text{em} = 0.0494 \pm 0.0009\\
    T_\text{COM} = 0.051 \pm 0.108\\
    T_\text{int} = 0.048 \pm 0.007\\

@@

    \label{eq:ekin}E_\text{Kin} &=& \frac{m}{2 M} \sum_i \vec{v}_i^{\, 2} \\
    \label{eq:temp}E &=& \frac{f}{2} k_B T

@@

    T = \frac{m}{3 M} \sum_i \vec{v}_i^{\, 2} \\

@@

Since this describes an instantaneous value, the mean value of the velocities over the course of the simulation has to be used to get the mean value
of the temperature of each subsystem. The simulation results compared with the mean values can be seen in

@@

The expected distribution is a gaussian with $\sigma=\sqrt{\frac{k_B T}{m}}$, i.e.

@@

    p(v_i) = \frac{m}{2\pi k_B T} e^{-\frac{x^2 m}{2 k_B T}}

@@

where $i$ denotes the chosen component of the center of mass velocity vector. A histogram of the x-component of the center of mass velocity  
with a curve representing \eqref{eq:gaussianvel} can be seen in Fig.~\ref{fig:comvel}. Since the y and z axis yield very similar histograms, the
x-component has been chosen to represent the velocity components. The graph in Fig.~\ref{fig:comvel} clearly shows, that the velocities indeed follow
the gaussian distribution.

@@

        $P=0.1$, $P=0.5$ and $P=1$ at impinging gas temperatures of $T_\text{imp}=0.05$ (red) and $T_\text{imp}=0.1$. The three columns each show 

@@

a value of a $T_\text{com} = 0.1$ at about $\Delta Q = 0.015$ at $P=0.1$, it takes a $\Delta Q$ of a little under $0.1$ to achieve the same value.

@@

correspont do a surrounding gas temperature of $T=0.1$) are growing slightly faster than the red curves ($T=0.05$).\\

>>>>

\lhead[\ ]{\leftmark }
\chead[\ ]{\ }
\rhead[\ ]{\rightmark }

\lfoot[\ ]{\ }
\cfoot[\ ]{\thepage }
\rfoot[\ ]{\ }

@@

\begin{document}
\title{Simulation of Nano Particles in a Laser Trap}
\author{Mathias H\"old, BSc.}
\date{2016}
\maketitle
\thispagestyle{empty}
\newpage
\tableofcontents
\newpage
\section{Introduction}
%something about computer science in general, optical tweezers and the need to study the
%systems on computers yada yada
In the year 1953, Metropolis et al.~\cite{Metropolis1953} used the MANIAC (Mathematical Analyzer Numerical Integrator And Computer) 
at the Los Alamos National Laboratories in the United States to
simulate particles in a liquid state to calculate the equation of state. This was the first publication in which scientists used a computer to
calculate numerical results based on a physical theory in a process that is to this day called \textit{Metropolis Monte Carlo simulation}. 
To say that since then a lot has
changed in the field of computer simulation (or computers in general) would be a vast understatement.\\
Not only has the processing power of computers increased exponentially (Moore's law), which allows the simulation of a greater number of particles for
a longer period of time, but also new techniques have been developed, such as molecular dynamics. These techniques are not only used by physicists,
but also find applications in the fields of chemistry, biology, etc., making it possible to apply them to virtually anything. In thhis thesis we 
apply methods of simulation to a field of active research: optical trapping.
Optical trapping is a technique that has been used and developed for over 40 years~\cite{Marago2013}. The idea for this goes back to the beginning of
the 20th century, where Lebedev~\cite{Lebedew1901}, Nichols and Hull~\cite{Nichols1901} demonstrated the existence of a force that light exerts on 
matter. Due to the technical limitations at that time, the idea was discarded until 1970, where Ashkin~\cite{Ashkin2000} used lasers to manipulate the
motion of neutral atoms and micrometer sized objects and thus brought the idea of optical trapping back to life.\\
With this advancement it was possible to develop methods for trapping and manipulating objects of different sizes ranging from the micrometer scale
down to the subnano scale. At first, this technique was used to either trap objects in the micrometer range (like cells) or in the subnano scale,
where lasers were used for the cooling of atoms. The range between those two, with objects at the scale of one to a hundred nanometer, was not
explored as much because neither downscaling the methods for micrometer sized objects nor upscaling the methods for atoms seemed feasible. In recent
years however, new developments in the field opened the possibilities to work with a variety of objects in the desired nanometer scale, such as metal
nanoparticles, nano tubes, quantum dots and much more. This non-invasive method of trapping objects allows very precise measurements of various
properties and also allows the cooling of nanoparticles towards their quantum mechanical ground state.\\
The small size of the objects used in the optical tweezer experiments make them amenable candidates for computer simulations. In this thesis, I will
model an optical tweezer setup, based on a real experiment, to investigate different temperatures in a non-equilibrium state.\\
The first chapter will be the introduction of the experiment and the introduction of the problem statement. The second chapter contains basic
information on the techniques used to model the experiment and simulate the dynamic properties. In the third chapter, I will present my findings which
are followed by conclusions and further remarks.

@@

The particle fluctuates within the trap in all three spatial directions. These fluctuations are approximatedly
decoupled, which means that they can be described by a one-dimensional Langevin equation:

@@

The one dimensional Langevin equation \eqref{eq:langevin} has to be modified to include the different friction coefficients of the two heat baths. The

@@

\newpage
\section{Simulation}
%simulation techniques and general concept of the simulation itself, used methods and so on 
The problem at hand can be studied on an atomic level with the use of computer simulation. There is a variety of methods for computer simulations
that are widely used, one of which being molecular dynamics (MD) simulations.\\ 
The goal is to simulate the experiment depicted in Fig.~\ref{fig:setup} as accurately as possible. To achieve this, the setup has to be broken down
into individual pieces that can be modeled by using existing methods. The nano particle will be approximated by a cube of particles, located on the
sites of a FCC
(face-centered cubic) lattice that interact via a Lennard-Jones potential. The laser will be broken down into its two main purposes, 
as it acts as a trapping device as well as a heat source for the nano particle. Surrounding the nano particle will be a gas chamber, 
that acts as a thermostat in the simulation, which will be modeled as a cubic box surrounding the nano particle. 
This box is filled with gas particles that interact with the nano particle via a soft-sphere potential and do not interact with one another.\\
The following section will give a brief overview of the concepts used to simulate every part of the experiment.

@@

Argon has a mass of $m$ = $6.69 \times 10^{-26}$ kg, $\sigma = 3.4 \times 10^{-10}$ m and $\varepsilon = 1.65 \times 10^{-21}$ J.\\

@@

The force in the $y$ and $z$ direction can be calculated analogously.\\

@@

energy after this point is vanishingly small. To save computing time it is common practice to cut off the potential at a certain point and not to

@@

    p_2 &=& \{0.5 \ a,0.5 \ a,0\}\\
    p_3 &=& \{0.5 \ a,0,0.5 \ a\}\\
    p_4 &=& \{0,0.5 \ a,0.5 \ a\}

@@

The sum of these two equations yields

@@

The problem with these kind of algorithms is that a target temperature has to be fixed which will be reached at some point. In order to simulate the

@@

Without any equilibrating mechanism the setup described by now would lead to the system heating up indefinitely, which is not desirable. Furthermore,

@@

    \item Randomly draw the number of particles that are created on a single side of the minimal volume of cells, $N_\text{fac}$, from a Poisson
        distribution with mean value

@@

        desired temperature. This chosen number of particles is then uniformly distributed over the face of the cell on which they are created. The
        velocities of the created particles are drawn from two different random number distributions. The component of the particle velocity perpendicular to

@@

        (where $U$ is the interaction energy between the gas particle and the crystal) because it is possible that the inserted particle is 

@@

    \mathbf{v}_\text{COM} &=& \frac1N\sum_{i=1}^N \mathbf{v}_i\\
    %\mathbf{v}_\text{cut} &=& \frac{\mathbf{v}_\text{tot}}N\\
    \mathbf{v}_i^\text{new} &=& \mathbf{v}_i - \mathbf{v}_\text{COM}

@@

The ideal gas thermostat/barostat, as described previously, 
keeps the system from overheating. The barostat algorithm is applied at the same time as the eHEX algorithm,

@@

The eHEX algorithm is pumping energy into the nano particle while the surrounding gas particles absorb excessive energy. This leads to the

@@

To check the integrity of the algortihm, a simple case needs to be tested. Such a simple case is $\Delta Q = 0$, i.e. no energy being pumped into the

@@

    T_\text{em} &=& 0.0494 \pm 0.0009\\
    T_\text{COM} &=& 0.051 \pm 0.108\\
    T_\text{int} &=& 0.048 \pm 0.007\\

@@

    \label{eq:ekin}E &=& \frac{m}{2 M} \sum_i \vec{v}_i^{\, 2} \\
    \label{eq:temp}E_\text{Kin} &=& \frac{f}{2} k_B T

@@

    k_B T = \frac{m}{3 M} \sum_i \vec{v}_i^{\, 2} \\

@@

Since this describes an instantaneous value, the mean value of the kinetic energy over the course of the simulation has to be used to get the mean value
of the temperature of each subsystem. The instantaneous values compared with the mean values can be seen in

@@

The expected distribution is a Gaussian with variance $\sigma^2=\frac{k_B T}{m}$, i.e.

@@

    p(v_i) = \sqrt{\frac{m}{2\pi k_B T}} e^{-\frac{x^2 m}{2 k_B T}}

@@

where $i$ denotes the chosen component of the center of mass velocity vector. A histogram of the $x$-component of the center of mass velocity  
with a curve representing eq.~\eqref{eq:gaussianvel} can be seen in Fig.~\ref{fig:comvel}. Since the $y$ and $z$ axis yield very similar histograms, the
$x$-component has been chosen to represent the velocity components. The graph in Fig.~\ref{fig:comvel} clearly shows, that the velocities indeed follow
the Gaussian distribution.

@@

            $P=0.1$, $P=0.5$ and $P=1$ at impinging gas temperatures of $T_\text{imp}=0.05$ (red) and $T_\text{imp}=0.1$ (blue). The three columns each show 

@@

a value of a $T_\text{com} = 0.1$ at about $\Delta Q = 0.015$ at $P=0.1$, it takes a $\Delta Q$ of a little under $0.1$ to achieve the same value
for $P = 1$.

@@

correspond to a surrounding gas temperature of $T=0.1$) are growing slightly faster than the red curves ($T=0.05$).\\
